Analytical Predictions of Explanet Obliquities Generated by Planet-Disk Interactions

### 2020-02-17 writeups/Su_disk_disp.tex
Forgot to check in work at home, this updates through end of S2

When $\eta < \eta_{\rm c}$, CS4 exists and is a saddle point. The two
trajectories originating and ending at CS4 are the only two infinite-period
orbits in the phase space. Together, these two critical trajectories are
referred to as the \emph{separatrix} and divide phase space into three zones. In
Eq.~\eqref{fig:eq_1contours}, we show the separatrix, the three zones, and their
relations to the CSs via a contour plot of the Hamiltonian of the system
\begin{align}
    \mathcal{H} &= -\frac{1}{2}\p{\hat{s} \cdot \hat{l}}^2
            + \eta \p{\hat{s} \cdot \hat{l}_{\rm d}},\label{eq:H}\\
        &= -\frac{1}{2}\cos^2\theta
            + \eta \p*{\cos \theta \cos I - \sin I \sin \theta \cos \phi}.
\end{align}
Note that trajectories in zone $II$ librate about CS2 while those in zones
$I$ and $III$ circulate.
\begin{figure*}
    \centering
    \includegraphics[width=0.8\textwidth]{../initial/0_eta/1contours_flip.png}
    \caption{Contour plot of $\mathcal{H}\p{\phi, \cos \theta}$ as given in
    Eq.~\eqref{eq:H}, where warmer colors denote more positive values. The black
    solid line is the separatrix, which only exists for $\eta < \eta_{\rm c}$.
    The three zones, divided by the separatrix, are labeled. The Cassini states
    are labeled and have the same colors as in Fig.~\ref{fig:cs_locs}. The
    interior of the separatrix, shaded in grey, is formally only defined for
    $\eta < \eta_{\rm c}$, but we may identify the points in phase space that
    flow into zone $II$ when evolved forward in time (decreasing $\eta$
    adiabatically); this is the shaded region in the top left panel, bounded by
    the black dotted line.}\label{fig:eq_1contours}
\end{figure*}

The unsigned areas $\p{\abs{\int \cos \theta \;\mathrm{d}\phi}}$ of the three
zones can be computed analytically \citep{henrard1987,ward2004I}. Defining
\begin{subequations}
    \begin{align}
        z_0 &= \eta\cos I, &
        \chi &= \sqrt{-\frac{\tan^3\theta_4}{\tan I} - 1},\\
        \rho &= \chi \frac{\sin^2 \theta_4\cos \theta_4}{
            \chi^2 \cos^2\theta_4 + 1},&
        T &= 2\chi \frac{\cos \theta_4}{
            \chi^2 \cos^2\theta_4 - 1},\\
    \end{align}
\end{subequations}
the areas for $\eta < \eta_{\rm c}$ are given by
\begin{subequations}\label{se:area_ward}
    \begin{align}
        A_{\rm II} &= 8\rho + 4\arctan T - 8z_0 \arctan \frac{1}{\chi},\\
        A_{\rm I} &= 2\pi\p{1 - z_0} - \frac{A_2}{2},\\
        A_{\rm III} &= 2\pi\p{1 + z_0} - \frac{A_2}{2}.
    \end{align}
\end{subequations}
These are plotted as a function of $\eta$ in Fig.~\ref{fig:eq_areas}. Note that
the zones are not formally defined for $\eta > \eta_{\rm c}$ since the
separatrix disappears, but a natural extension exists: evolve an initial phase
space point $p$ under adiabatic decrease of $\eta$ until the separatrix appears
at $\eta = \eta_{\rm c}$, then identify $p$ with the zone it is in at $\eta_{\rm
c}$. Since phase space area is conserved under adiabatic evolution, this implies
$A_{\rm i}\p{\eta > \eta_{\rm c}} = A_{\rm i}(\eta_{\rm c})$. The boundary between these
extended zones is denoted by the dashed black line in the top left panel of
Fig.~\ref{fig:eq_1contours}, where otherwise no separatrix exists.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/99_misc/1_areas.png}
    \caption{Plot of fractional areas of each of the zones $A_{\rm i}(\eta) /
    4\pi$ as given by Eqs.~\eqref{se:area_ward}. Dotted lines correspond to
    small $\eta$ approximations discussed in Appendix~\ref{s:ad_approx}. Dashed
    lines for $\eta > \eta_{\rm c}$ effective values of $A_{\rm II}, A_{\rm
    III}$ for $\eta > \eta_{\rm c}$, denoting the area of points that would flow
    into either area under adiabatic decrease of $\eta$ from $\eta > \eta_{\rm
    c}$.}\label{fig:eq_areas}
\end{figure}

@@

As a particular example, consider a system in zone $II$ in panel (d) of
Fig.~\ref{fig:eq_1contours}. As $\eta$ decreases, zone $II$ will shrink while
zones $I$ and $III$ will expand until the trajectory. Suppose the trajectory
exits zone $II$ at some $\eta_\star$, then the probability of a $II \to I$
transition is $\Pr\p{II \to I} = -\frac{\dot{A}_{\rm I}}{\dot{A}_{\rm II}}$,
while a $II \to III$ transition occurs with probability $\Pr\p{II \to III} =
-\frac{\dot{A}_{\rm III}}{\dot{A}_{\rm II}}$.

@@

Returning to the evolution of $\hat{s}$, we can classify trajectories by the
separatrix encounters experienced. Initially, in the $\eta > \eta_{\rm c}$ regime,
only zones $II, III$ exist. Conversely, at the end of the simulation when $\eta
\to 0$, only zones $I, III$ exist. Most simply, one might expect four sequences
of transitions between zones (we call these dynamical ``tracks'') to manifest:
from one of $\z{I, III}$ to one of $\z{II, III}$, where $III \to III$ is a
trivial case (no separatrix encounter). In addition to these four tracks, a
fifth track $III \to II \to I$ is observed. Below, we will describe the
evolution of $A$ throughout each of the five tracks, as well as determine the
initial conditions and probabilities associated with each track:
\begin{enumerate}
    \item $II \to I$ --- An example trajectory following this track is depicted
        in Fig.~\ref{fig:ad_21}. $\hat{s}$ starts librating about CS2 in zone
        $II$, enclosing some initial phase space area $A_{\rm i}$. The trajectory is
        able to librate without separatrix encounter until $A_{\rm
        II}(\eta_\star) = A_{\rm i}$. As the trajectory transitions to a circulating
        trajectory in zone $I$ immediately bordering the separatrix, it will
        encompass $-A_{\rm I}(\eta_\star)$ phase space area. The final
        $\theta_{\rm f}$ is then given by Eq.~\eqref{eq:qfaf}. An analytical
        approximation to $\theta_{\rm f}$ is derived in
        Appendix~\ref{s:ad_approx} and is
        \begin{equation}
            \cos \theta_{\rm f, II \Rightarrow I} \approx
                \p{\frac{\pi \theta_{\rm sp, i}^2}{16}}^2 \cot I
                    + \frac{\theta_{\rm sp, i}^2}{4}.\label{eq:qf_21_approx}
        \end{equation}

        This transition can only occur when the initial condition begins in zone
        II, requiring $A_{\rm i} < A_{\rm II}(\eta_{\rm c})$. Note $A_{\rm II}(\eta_{\rm c})$ is
        given in closed form in \citet{ward2004I} but appears to be incorrect.
        Then, at some known $\eta_\star$ satisfying $A_{\rm II}(\eta_\star) =
        A_{\rm i}$, the trajectory is ejected from the separatrix following
        Eq.~\eqref{eq:henrard_hop}. Note that since $\pd{A_{\rm i}}{\eta} < 0$
        everywhere, while $\pd{A_{\rm II}}{\eta} > 0$ at all possible
        $\eta_\star$ for an initial condition starting in zone $II$, this track
        always has nonzero probability.

    \item $II \to III$ --- An example trajectory following this track is
        depicted in Fig.~\ref{fig:ad_23}. The only difference from the previous
        track is that, upon separatrix encounter, the trajectory follows the
        circulating trajectory in zone $III$ bordering the separatrix, upon
        which it will encompass $A_{\rm f} = A_{\rm I}(\eta_\star) + A_{\rm
        II}(\eta_\star)$. The final obliquity is still given by
        Eq.~\eqref{eq:qfaf}, and the analytical approximation derived in
        Appendix~\ref{s:ad_approx} is
        \begin{equation}
            \cos \theta_{\rm f, II \Rightarrow III} \approx
                \p{\frac{\pi \theta_{\rm sp, i}^2}{16}}^2 \cot I
                    - \frac{\theta_{\rm sp, i}^2}{4}.\label{eq:qf_23_approx}
        \end{equation}

        Again, this track can only occur when $A_{\rm i} < A_{\rm II}(\eta_{\rm c})$, but a
        further constraint arises when we consider the transition probability
        given by Eq.~\eqref{eq:henrard_hop}. Upon examination of
        Fig.~\ref{fig:eq_areas}, it is clear that $\pd{A_{\rm III}}{\eta} > 0$ for
        many $\eta$. Call
        \begin{equation}
            \eta_{\min, III} \equiv \argmin A_{\rm III}(\eta),
        \end{equation}
        which is labeled in Fig.~\ref{fig:eq_areas}, then if $\eta_\star >
        \eta_{\rm \min, III}$ then $\Pr_{\rm II \to III} < 0$. This is understood as a
        forbidden transition, and so $II \to III$ is only a permitted dynamical
        track if $\eta_\star <, \eta_{\rm \min, III}$.

    \item $III \to I$ --- This track is depicted in Fig.~\ref{fig:ad_31}.
        The trajectory encounters the separatrix when $A_{\rm I}(\eta_\star) +
        A_{\rm II}(\eta_\star) = A_{\rm i}$, upon which it transitions to a zone $I$
        trajectory enclosing $A_{\rm f} = -A_{\rm I}$. Then, as always, the final
        obliquity is given by Eq.~\eqref{eq:qfaf}.

        This track can only occur if $A_{\rm i} > A_{\rm II}(\eta_{\rm c})$, but is also
        constrained by requiring its $A_{\rm i}$ is sufficiently small it will
        encounter the separatrix (if it is too large, it will never encounter
        the separatrix, which is a $III \to III$ transition). This is written
        $A_{\rm i} < \max A_{\rm I} + A_{\rm II} = 4\pi - \min A_{\rm III}$.

@@

    \item $III \to II \to I$ --- This track is depicted in Fig.~\ref{fig:ad_321}.
        The first separatrix encounter occurs at $\eta_1$ when $A_{\rm
        I}(\eta_1) + A_{\rm II}(\eta_1) = A_{\rm i}$, upon which the trajectory moves
        into zone $II$ enclosing intermediate phase space area $A_{\rm m} = A_{\rm
        II}(\eta_1)$. Examining Fig.~\ref{fig:eq_areas}, there exist values $A$
        for which $A_{\rm II}(\eta) = A$ has multiple solutions, and $\eta_1 >
        \eta_{\rm II, \max}$ as labeled. These are the only values for which the
        $III \to II \to I$ transition is permitted. Thus, as $\eta$ continues to
        decrease, a second $\eta_2$ value exists for which $A_{\rm m} = A_{\rm
        II}(\eta_2)$, upon which the trajectory is ejected to zone $I$ and $A_f
        = -A_{\rm I}(\eta_2)$. The final obliquity again then obeys
        Eq.~\eqref{eq:qfaf}.

        Similar to the $III \to I$ track, the $III \to II \to I$ track can only
        occur over the same $A_{\rm II}(\eta_{\rm c}) < A_{\rm i} < \max A_{\rm I} + A_{\rm
        II}$ interval. The probability of encountering this track is set by the
        initial $III \to II$ transition. It bears noting that this probability
        is only nonnegative for a very small fraction of $A_{\rm i}$ values, since it
        requires $\pd{A_{\rm II}}{\eta}$ and $\pd{A_{\rm III}}{\eta}$ to have
        different signs; this occurs only if $A_{\rm II}\p{\eta_{\rm c}} <
        A_{\rm i} <
        A_{\rm II, \max}$. Outside of these bounds, $\Pr_{\rm III \to II} < 0$
        which is interpreted again as a forbidden transition.

        Then, once a $III \to II$ transition occurs, the second $II \to I$
        transition occurs for some $\eta_2$ satisfying $A_{\rm II}(\eta_2) =
        A_{\rm II}(\eta_1), \eta_2 < \eta_1$. Graphical inspection shows that
        $\pd{A_{\rm II}}{\eta}$ has the same sign as $\pd{A_{\rm III}}{\eta}$
        over all possible values, so the second $II \to I$ transition is
        guaranteed, completing the $III \to II \to I$ track.

    \item $III \to III$ --- This track depicts the trivial case where no
        separatrix encounter ever occurs, and $A$ is constant throughout the
        evolution. This constraint equates to $A_{\rm i} > \max A_{\rm I} + A_{\rm
        II}$, whereupon no separatrix encounter occurs at all, and $A_{\rm f} =
        A_{\rm i}$,
        which for $\eta_{\rm i} \to \infty, \eta_{\rm f} \to 0$ results in $\theta =
        \theta_{\rm sd, i}$.

        Note that a small deviation from $\theta = \theta_{\rm sd, i}$ occurs
        because we use finite $\eta_{\rm i}$, as hinted at in Section~\ref{sss:a_evo}.
        This is because for $\theta_{\rm sd, i} \gtrsim \pi/2$, trajectories are
        better described as librating about CS3. Nevertheless, we can still
        approximate $A_{\rm i}$ to good accuracy as the solid angle enclosed by
        libration about CS3 with polar angle $2\pi + \theta_3 - \theta_2 -
        \theta_{\rm sd, i}$. This recovers the correct transfer function
        $\theta_{\rm f}\p{\theta_{\rm sd, i}}$ for the $III \to III$ regime, as
        can be seen visibly in Fig.~\ref{fig:3_ensemble_20_35}.
\end{enumerate}
In summary, the five regimes of track possibilities in increasing order of
$A_{\rm i}$
are:
\begin{enumerate}
    \item $A_{\rm i} \in \s{0, A_{\rm II}\p{\eta_{\rm \min, III}}}$ --- $II \to III,
        II \to I$ both possible.

    \item $A_{\rm i} \in \s{A_{\rm II}\p{\eta_{\rm \min, III}}, A_{\rm II}(\eta_{\rm c})}$
        --- $II \to I$ only.

    \item $A_{\rm i} \in \s{A_{\rm II}(\eta_{\rm c}), A_{\rm II, \max}}$ --- $III \to I, III
        \to II \to I$ both possible.

    \item $A_{\rm i} \in \s{A_{\rm II, \max}, \max A_{\rm I} + A_{\rm II}}$ --- $III
        \to I$ only.

>>>>

The Hamiltonian (in the rotating frame) of the system is
\begin{align}
    \mathcal{H} &= -\frac{1}{2}\p{\hat{s} \cdot \hat{l}}^2
            + \eta \p{\hat{s} \cdot \hat{l}_{\rm d}}\label{eq:H}\nonumber\\
        &= -\frac{1}{2}\cos^2\theta
            + \eta \p{\cos \theta \cos I - \sin I \sin \theta \cos \phi}.
\end{align}
Trajectories in the phase space $\p{\phi, \cos \theta}$ satisfy $H = $ constant
(see Fig.~\autoref{fig:eq_1contours}).

When $\eta < \eta_{\rm c}$, CS4 exists and is a saddle point. The two
trajectories originating and ending at CS4 are the only two infinite-period
orbits in the phase space. Together, these two critical trajectories are
referred to as the \emph{separatrix} and divide phase space into three zones. In
Fig.~\ref{fig:eq_1contours}, we show the separatrix, the three zones, and their
relations to the CSs. Note that trajectories in zone II librate about CS2
while those in zones I and III circulate.
\begin{figure*}
    \centering
    \includegraphics[width=0.8\textwidth]{../initial/0_eta/1contours_flip.png}
    \caption{Level curves plot of $\mathcal{H}\p{\phi, \cos \theta}$
    (Eq.~\eqref{eq:H}) for $I = 5^\circ$, where warmer colors denote more
    positive values. The black solid line is the separatrix, which only exists
    for $\eta < \eta_{\rm c} = 0.766$. The three zones (I, II, III), divided by
    the separatrix, are labeled. The Cassini states are denoted by filled
    circles and have the same colors as in Fig.~\ref{fig:cs_locs}. The interior
    of the separatrix, shaded in grey, is formally only defined for $\eta <
    \eta_{\rm c}$, but we may identify the points in phase space that flow into
    zone II when evolved forward in time (decreasing $\eta$ adiabatically); this
    is the shaded region in the top left panel, bounded by the black dotted
    line.}\label{fig:eq_1contours}
\end{figure*}

Since $\p*{\phi, \cos \theta}$ form a set of canonical variables, the integral
$\oint \cos \theta\;\mathrm{d}\phi$ along a trajectory is an adiabatic invariant
(see Section~\autoref{s:ad}). The unsigned areas $\p{\abs{\int \cos \theta
\;\mathrm{d}\phi}}$ of the three zones (as defined in
Fig.~\autoref{fig:eq_1contours}) can be computed analytically
\citep{henrard1987,ward2004I}. Define
\begin{subequations}
    \begin{align}
        z_0 &= \eta\cos I, &
        \chi &= \sqrt{-\frac{\tan^3\theta_4}{\tan I} - 1},\\
        \rho &= \chi \frac{\sin^2 \theta_4\cos \theta_4}{
            \chi^2 \cos^2\theta_4 + 1},&
        T &= 2\chi \frac{\cos \theta_4}{
            \chi^2 \cos^2\theta_4 - 1}.
    \end{align}
\end{subequations}
The areas for $\eta < \eta_{\rm c}$ are given by
\begin{subequations}\label{se:area_ward}
    \begin{align}
        A_{\rm I} &= 2\pi\p{1 - z_0} - \frac{A_2}{2},\\
        A_{\rm II} &= 8\rho + 4\arctan T - 8z_0 \arctan \frac{1}{\chi},\\
        A_{\rm III} &= 2\pi\p{1 + z_0} - \frac{A_2}{2}.
    \end{align}
\end{subequations}
These are plotted as a function of $\eta$ in Fig.~\ref{fig:eq_areas}. Note that
the zones are not formally defined for $\eta > \eta_{\rm c}$ since the
separatrix disappears, but a natural extension exists: evolve an initial phase
space point $p$ under adiabatic decrease of $\eta$ until the separatrix appears
at $\eta = \eta_{\rm c}$, then identify $p$ with the zone it is in at $\eta_{\rm
c}$. Since phase space area is conserved under adiabatic evolution, this
extension implies $A_{\rm i}\p{\eta > \eta_{\rm c}} = A_{\rm i}(\eta_{\rm c})$.
The boundary between these extended zones is denoted by the dashed black line in
the top left panel of Fig.~\ref{fig:eq_1contours}, where otherwise no separatrix
exists.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/99_misc/1_areas.png}
    \caption{Fractional areas of each of the zones $A_{\rm i}(\eta) /
    4\pi$ as given by Eqs.~\eqref{se:area_ward}. The dotted lines correspond to
    small $\eta$ approximations discussed in Appendix~\ref{s:ad_approx}. The
    dashed lines for $\eta > \eta_{\rm c}$ are the effective values of $A_{\rm
    II}, A_{\rm III}$ for $\eta > \eta_{\rm c}$, denoting the points that would
    flow into either area under adiabatic decrease of $\eta$ from $\eta >
    \eta_{\rm c}$. The vertical lines correspond to $\eta = \eta_{\rm c}$
    (Eq.\eqref{eq:etac}) and the values of $\eta$ for which $A_{\rm II}$ is
    maximized ($\eta_{\rm \max, II}$) and for which $A_{\rm III}$ is minimized
    ($\eta_{\rm \min, III}$, Eq.~\eqref{eq:eta_minIII}).}\label{fig:eq_areas}
\end{figure}

@@

As a particular example, consider a system in zone II in panel (d) of
Fig.~\ref{fig:eq_1contours}. As $\eta decreases, zone II$ will shrink while
zones I and III will expand until the trajectory. Suppose the trajectory
exits zone II at some $\eta_\star$, then the probability of a II $\to$ I
transition is $\Pr\p{\rm II \to I} = -\frac{\dot{A}_{\rm I}}{\dot{A}_{\rm II}}$,
while a II $\to$ III transition occurs with probability $\Pr\p{\rm II \to III} =
-\frac{\dot{A}_{\rm III}}{\dot{A}_{\rm II}}$.

@@

Returning to the evolution of $\hat{s}$, we can classify trajectories by the
separatrix encounters experienced. Initially, in the $\eta > \eta_{\rm c}$ regime,
only zones II and III exist. Conversely, at the end of the simulation when $\eta
\to 0$, only zones I and III exist. Most simply, one might expect four sequences
of transitions between zones (we call these dynamical ``tracks'') to manifest:
from one of $\z{I, III}$ to one of $\z{II, III}$, where III $\to$ III is a
trivial case (no separatrix encounter). In addition to these four tracks, a
fifth track III $\to$ II $\to$ I is observed. Below, we will describe the
evolution of $A$ throughout each of the five tracks, as well as determine the
initial conditions and probabilities associated with each track:
\begin{enumerate}
    \item II $\to$ I --- An example trajectory following this track is depicted
        in Fig.~\ref{fig:ad_21}. $\hat{s}$ starts librating about CS2 in zone
        II, enclosing some initial phase space area $A_{\rm i}$. The trajectory is
        able to librate without separatrix encounter until $A_{\rm
        II}(\eta_\star) = A_{\rm i}$. As the trajectory transitions to a circulating
        trajectory in zone I immediately bordering the separatrix, it will
        encompass $-A_{\rm I}(\eta_\star)$ phase space area. The final
        $\theta_{\rm f}$ is then given by Eq.~\eqref{eq:qfaf}. An analytical
        approximation to $\theta_{\rm f}$ is derived in
        Appendix~\ref{s:ad_approx} and is
        \begin{equation}
            \cos \theta_{\rm f, II \Rightarrow I} \approx
                \p{\frac{\pi \theta_{\rm sp, i}^2}{16}}^2 \cot I
                    + \frac{\theta_{\rm sp, i}^2}{4}.\label{eq:qf_21_approx}
        \end{equation}

        This transition can only occur when the initial condition begins in zone
        II, requiring $A_{\rm i} < A_{\rm II}(\eta_{\rm c})$. Note $A_{\rm II}(\eta_{\rm c})$ is
        given in closed form in \citet{ward2004I} but appears to be incorrect.
        Then, at some known $\eta_\star$ satisfying $A_{\rm II}(\eta_\star) =
        A_{\rm i}$, the trajectory is ejected from the separatrix following
        Eq.~\eqref{eq:henrard_hop}. Note that since $\pd{A_{\rm i}}{\eta} < 0$
        everywhere, while $\pd{A_{\rm II}}{\eta} > 0$ at all possible
        $\eta_\star$ for an initial condition starting in zone II, this track
        always has nonzero probability.

    \item II $\to$ III --- An example trajectory following this track is
        depicted in Fig.~\ref{fig:ad_23}. The only difference from the previous
        track is that, upon separatrix encounter, the trajectory follows the
        circulating trajectory in zone III bordering the separatrix, upon
        which it will encompass $A_{\rm f} = A_{\rm I}(\eta_\star) + A_{\rm
        II}(\eta_\star)$. The final obliquity is still given by
        Eq.~\eqref{eq:qfaf}, and the analytical approximation derived in
        Appendix~\ref{s:ad_approx} is
        \begin{equation}
            \cos \theta_{\rm f, II \Rightarrow III} \approx
                \p{\frac{\pi \theta_{\rm sp, i}^2}{16}}^2 \cot I
                    - \frac{\theta_{\rm sp, i}^2}{4}.\label{eq:qf_23_approx}
        \end{equation}

        Again, this track can only occur when $A_{\rm i} < A_{\rm II}(\eta_{\rm c})$, but a
        further constraint arises when we consider the transition probability
        given by Eq.~\eqref{eq:henrard_hop}. Upon examination of
        Fig.~\ref{fig:eq_areas}, it is clear that $\pd{A_{\rm III}}{\eta} > 0$ for
        many $\eta$. Call
        \begin{equation}
            \eta_{\min, III} \equiv \argmin A_{\rm III}(\eta)
                \label{eq:eta_minIII},
        \end{equation}
        which is labeled in Fig.~\ref{fig:eq_areas}, then if $\eta_\star >
        \eta_{\rm \min, III}$ then $\Pr_{\rm II \to III} < 0$. This is
        understood as a forbidden transition, and so II $\to$ III is only a
        permitted dynamical track if $\eta_\star <, \eta_{\rm \min, III}$.

    \item III $\to$ I --- This track is depicted in Fig.~\ref{fig:ad_31}.
        The trajectory encounters the separatrix when $A_{\rm I}(\eta_\star) +
        A_{\rm II}(\eta_\star) = A_{\rm i}$, upon which it transitions to a zone
        I trajectory enclosing $A_{\rm f} = -A_{\rm I}$. Then, as always, the
        final obliquity is given by Eq.~\eqref{eq:qfaf}.

        This track can only occur if $A_{\rm i} > A_{\rm II}(\eta_{\rm c})$, but is also
        constrained by requiring its $A_{\rm i}$ is sufficiently small it will
        encounter the separatrix (if it is too large, it will never encounter
        the separatrix, which is a III $\to$ III transition). This is written
        $A_{\rm i} < \max A_{\rm I} + A_{\rm II} = 4\pi - \min A_{\rm III}$.

@@

    \item III $\to$ II $\to$ I --- This track is depicted in
        Fig.~\ref{fig:ad_321}. The first separatrix encounter occurs at $\eta_1$
        when $A_{\rm I}(\eta_1) + A_{\rm II}(\eta_1) = A_{\rm i}$, upon which
        the trajectory moves into zone II enclosing intermediate phase space
        area $A_{\rm m} = A_{\rm II}(\eta_1)$. Examining
        Fig.~\ref{fig:eq_areas}, there exist values $A$ for which $A_{\rm
        II}(\eta) = A$ has multiple solutions, and $\eta_1 > \eta_{\rm II,
        \max}$ as labeled. These are the only values for which the III $\to$ II
        $\to$ I transition is permitted. Thus, as $\eta$ continues to decrease,
        a second $\eta_2$ value exists for which $A_{\rm m} = A_{\rm
        II}(\eta_2)$, upon which the trajectory is ejected to zone I and $A_f =
        -A_{\rm I}(\eta_2)$. The final obliquity again then obeys
        Eq.~\eqref{eq:qfaf}.

        Similar to the III $\to$ I track, the III $\to$ II $\to$ I track can
        only occur over the same $A_{\rm II}(\eta_{\rm c}) < A_{\rm i} < \max
        A_{\rm I} + A_{\rm II}$ interval. The probability of encountering this
        track is set by the initial III $\to$ II transition. It bears noting
        that this probability is only nonnegative for a very small fraction of
        $A_{\rm i}$ values, since it requires $\pd{A_{\rm II}}{\eta}$ and
        $\pd{A_{\rm III}}{\eta}$ to have different signs; this occurs only if
        $A_{\rm II}\p{\eta_{\rm c}} < A_{\rm i} < A_{\rm II, \max}$. Outside of
        these bounds, $\Pr_{\rm III \to II} < 0$ which is interpreted again as a
        forbidden transition.

        Then, once a III $\to$ II transition occurs, the second II $\to$ I
        transition occurs for some $\eta_2$ satisfying $A_{\rm II}(\eta_2) =
        A_{\rm II}(\eta_1), \eta_2 < \eta_1$. Graphical inspection shows that
        $\pd{A_{\rm II}}{\eta}$ has the same sign as $\pd{A_{\rm III}}{\eta}$
        over all possible values, so the second II $\to$ I transition is
        guaranteed, completing the III $\to$ II $\to$ I track.

    \item III $\to$ III --- This track depicts the trivial case where no
        separatrix encounter ever occurs, and $A$ is constant throughout the
        evolution. This constraint equates to $A_{\rm i} > \max A_{\rm I} + A_{\rm
        II}$, whereupon no separatrix encounter occurs at all, and $A_{\rm f} =
        A_{\rm i}$,
        which for $\eta_{\rm i} \to \infty, \eta_{\rm f} \to 0$ results in $\theta =
        \theta_{\rm sd, i}$.

        Note that a small deviation from $\theta = \theta_{\rm sd, i}$ occurs
        because we use finite $\eta_{\rm i}$, as hinted at in Section~\ref{sss:a_evo}.
        This is because for $\theta_{\rm sd, i} \gtrsim \pi/2$, trajectories are
        better described as librating about CS3. Nevertheless, we can still
        approximate $A_{\rm i}$ to good accuracy as the solid angle enclosed by
        libration about CS3 with polar angle $2\pi + \theta_3 - \theta_2 -
        \theta_{\rm sd, i}$. This recovers the correct transfer function
        $\theta_{\rm f}\p{\theta_{\rm sd, i}}$ for the III $\to$ III regime, as
        can be seen visibly in Fig.~\ref{fig:3_ensemble_20_35}.
\end{enumerate}
In summary, the five regimes of track possibilities in increasing order of
$A_{\rm i}$
are:
\begin{enumerate}
    \item $A_{\rm i} \in \s{0, A_{\rm II}\p{\eta_{\rm \min, III}}}$ --- II $\to$
        III and II $\to$ I both possible.

    \item $A_{\rm i} \in \s{A_{\rm II}\p{\eta_{\rm \min, III}}, A_{\rm
        II}(\eta_{\rm c})}$ --- II $\to$ I only.

    \item $A_{\rm i} \in \s{A_{\rm II}(\eta_{\rm c}), A_{\rm II, \max}}$ ---
        III $\to$ I and III $\to$ II $\to$ I both possible.

    \item $A_{\rm i} \in \s{A_{\rm II, \max}, \max A_{\rm I} + A_{\rm II}}$ ---
        III $\to$ I only.

    \item $A_{\rm i} > \max A_{\rm I} + A_{\rm II}$ --- III $\to$ III only.
\end{enumerate}
The corresponding ranges for for $\theta_{\rm sd,i}$ can be computed via
Eq.~\eqref{eq:ai_qsd}. The boundaries between these ranges are overplotted in
Fig.~\ref{fig:ad_ensemble}, where they can be seen to agree well with numerical
results.
\begin{figure}
    \centering
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3testo21.png}
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3testo21_subplots.png}
    \end{subfigure}
    \caption{Fiducial simulation following the $A_{\rm II} \to A_{\rm I}$
    transition. An initial $\theta_{\rm sd, i} = 0.3\;\mathrm{rad} \approx
    17.2^\circ$ was used, as well as $\epsilon = 3 \times 10^{-4}$. Top plot
    upper panel: plot of $\cos \theta(t)$ (green) in an example simulation.
    Overlaid are the locations of Cassini State 2 (red), and upper and lower
    bounds on the separatrix (black). Note that the trajectory successfully
    tracks CS2 to an obliquity $\theta \approx 90^\circ$. Light vertical blue
    lines denote times portrayed in bottom plot. Top plot lower panel: plot of
    the enclosed separatrix area obtained by integrating the simulated
    trajectory (blue dots) and adiabatic theory (red line). Even though the
    enclosed area jumps at separatrix crossing, the resultant $A$ is still
    accurately predicted. Lower plot: snapshots in the $\p{\cos \theta, \phi}$
    space of the simulation trajectories for the vertical blue lines in the top
    plots. These times correspond to the start of the simulation, the appearance
    of the separatrix, two panels depicting the separatrix crossing process, and
    a final snapshot at $\eta = 10^{-3.5}$. The green line denotes a full
    circulation or libration cycle at the selected time, including an arrow for
    directionality. The separatrix at the start/end of the portrayed cycle are
    shown in solid/dashed black lines respectively. Also labeled is CS2 at the
    start of the cycle (red). Finally, the enclosed phase space area is shaded
    in grey ($A > 0$) and red ($A < 0$).}\label{fig:ad_21}
\end{figure}
\begin{figure}
    \centering
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3testo23.png}
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3testo23_subplots.png}
    \end{subfigure}
    \caption{Same as Fig.~\ref{fig:ad_21} but for the $A_2 \to A_3$ track.
    $\theta_{\rm sd, i} = 0.3\;\mathrm{rad}$ and $\epsilon = 3.01 \times
    10^{-4}$.}\label{fig:ad_23}
\end{figure}
\begin{figure}
    \centering
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3testo31.png}
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3testo31_subplots.png}
    \end{subfigure}
    \caption{Same as Fig.~\ref{fig:ad_21} but for the $A_{\rm III} \to A_{\rm I}$
    track. $\theta_{\rm sd, i} = 89.1^\circ$, and $\epsilon = 3 \times
    10^{-4}$.}\label{fig:ad_31}
\end{figure}
\begin{figure}
    \centering
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3testo321.png}
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3testo321_subplots.png}
    \end{subfigure}
    \caption{Same as Fig.~\ref{fig:ad_21} but for the $A_{\rm III} \to A_{\rm II}
    \to A_{\rm I}$ track. $\theta_{\rm sd, i} = 60^\circ$, and $\epsilon = 3.14
    \times 10^{-4}$. Two separatrix crossings are shown.}\label{fig:ad_321}
\end{figure}
